% Section about analog MC and the challenges faced when trying to calculate SDR in
% FES with no VR.

\subsection{Analog Monte Carlo Calculations}

To quantify the dose rate, we need to know th flux of photons
produced by the decaying radionuclides in the activated components. The
activation is causedd by neutron irradiation.
We need to know what the neutron flux looks like in every region of phase
space.

We need a detailed distribution of the neutron flux throughout all regions of phase
space.
The neutron distribution along with and irradiation and decay scheme of interest 
can then be input for a nuclear inventory code to generate the photon
emission density.  This photon emission density is used as a source for the
photon transport step which ultimately produces the photon flux in every region.
The photon flux coupled with a flux-to-dose-rate conversion factor will give us
the dose rate.


There are two ways to solve the Boltzmann transport equation: deterministically
and stochastically.
The most optimal way to perform accurate, full-scale analysis of FES is
through Monte Carlo radiation transport *need source*.
In general, Monte Carlo (MC) calculations rely on repeated, random sampling to solve
mathematical problems.  The MC method can be applied to radiation transport by
solving the Boltzmann transport equation through the simulation of random particle
walks through phase space.  In analog operation mode (i.e. no variance reduction), 
the source particle's position, energy, direction
and subsequent collisions are sampled from probability
distribution functions (PDFs).  Quantities of interest such as flux can be
scored, or tallied, by averaging particle behavior
in discrete regions of phase space.

Radiation transport simulations in geometries that have thick shielding, such as
FES, become challenging for MC codes.  The particles have lots of interactions
(absorption and scattering) in the shielding regions.  This results in low
particle fluxes in these discrete regions of phase space.  

The statistical error is a function of the relative error, R, which is defined as
\begin{equation} \label{eq:2}
		R = \frac{S_{\overline{x}}}{{\overline{x}}}
\end{equation}
where $\overline{x}$ is the average of the tally scores, and $S_{\overline{x}}$ is the standard
deviation of the tally scores.  For a well behaved tally, R is proportional
to $1/\sqrt{N}$ where N is the number of tally scores \cite{mcnp_manual}.

To reliably predict results in these regions, many particle
histories need to be simulated which may require large amounts of computer time
\cite{haghighat_wagner_2003}.

MC calculation efficiency is measured by a quantity known as the figure of merit
(FOM).  The FOM is a function of relative error, R, and computer processing
time, T, 
as given by
\begin{equation} \label{eq:3}
		FOM = \frac{1}{{{R^2}T}}
\end{equation}
It is desirable to have a high FOM because it means that less computation time is needed to achieve
a reasonably low error, less than 0.1 according to the MCNP manual
\cite{mcnp_manual}. 
The relative error is inversely proportional to N.
VR methods
aim to increase the FOM by increasing N and decreasing $S_{\overline{x}}$ by sampling from biased PDFs; this effectively
forces more collisions in regions of phase space that are
important to the tally of interest. 
